Réponse et Résilience de la Biodiversité d'une Forêt Tropicale après Perturbation

### 2018-09-12 04-TrajectoiresDeDiversité.Rmd
discussion_commBH

Dans ce chapitre nous nous intéresserons aux trajectoires de diversité et de composition à l'échelle de l'ensemble de la communauté.
La diversité des forêts tropicales est supposée être soumise et entretenue naturellement par un régime régulier de perturbations, qui entraîne un pic de diversité pour des perturbations d'intensité et de de fréquence moyennes.
En milieu tropical cependant cette théorie des perturbation intermédiaires reste débattue, ce qui questionne la résilience fonctionnelle et taxonomique des forêts tropicales.

@@

Les trajectoires de diversité ont mis en évidence des réponses taxonomique et fonctionnelle cycliques, permettant la restauration des caractéristiques d'avant perturbation.
Les trajectoires ont montré la divergence taxonomique des communautés, due aux limites de dispersion des espèces maintenant les différences de composition initiales, et leur convergence fonctionnelle, due à des processus déterministes imposant une trajectoire fonctionnelle commune.
La théorie des perturbations intermédiaires a montré prédire les trajectoires de diversité taxonomique, augmentant avec l'intensité de la perturbation jusqu'à un certain seuil.
En revanche la diversité fonctionnelle augmentaient après perturbation quelle qu'en soit l'intensité, l'impact de la perturbation étant atténué par l'importante redondance fonctionnelle des communautés.
Bien qu'effective, la restauration des caractéristiques taxonomiques et fonctionnelles des communautés restait inachevée après 30 ans, suggérant un temps de restauration long de plusieurs décennies d'autant plus difficile à estimer que la restauration de la redondance fonctionnelle, déterminante de la résilience des communautés, s'est montrée ralentir au cours du temps.

@@

Ariane MIRABEL ^1^ ^\*^, Bruno Hérault ^2^, Eric Marcon ^1^
\newline

@@

Validations of the IDH though remain scarce in the long-term and mainly rely on the analysis of taxonomic richness [@Molino2001]. Taxonomic richness alone, however, gives limited or misleading information on forests recovery and functioning [@Chaudhary2016]. More ecological-meaningful analysis would couple richness with (i) evenness, that would reveal the changes in the species abundance distribution and thus the underlying ecological processes, and (ii) composition that is crucial for conservation issues [@Lavorel2002;@Bellwood2006]. Furthermore, a functional approach accounting for species biological attributes would directly link communities diversity, composition and redundancy to ecosystem functioning and to its environmental constraints [@Violle2007b;@Baraloto2012a].
In that respect, the functional trait-based approach that focus on major traits related to species ecology and mediate species performance in a given environment was successfully adopted [@Diaz2005].
For instance, the functional approach revealed in tropical rainforests the deterministic processes entailing, after disturbance, a functional shift from a dominance of "conservative" slow-growing species dealing with scarce resources to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@Reich2014;@Herault2011].
This shift is translated into the trajectories of key functional traits related to resource acquisition (leaf and stem traits) and life-history traits (seed mass, maximum size) [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
Eventually a complete overview of community response to disturbance would encompass the changes in functional redundancy, that quantifies the amount of shared trait values among species [@Carmona2016].
The high functional redundancy of hyper-diverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines the resilience of communities after disturbance [@Elmqvist2003; @Diaz2005].

@@

The IDH well predicted well the disturbance impact on community taxonomic richness, enhanced until an intensity threshold (20-25% AGB loss), and to some extent on taxonomic evenness, somewhat decoupled from the disturbance intensity as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
The disturbance intensity determined the balance in the community between pre-disturbance surviving trees and those recruited afterward.
The pool of true pioneer species specifically recruited after disturbance is restricted in the Guiana Shield to a few common genera (e.g. Cecropia spp., Vismia spp.) [@Guitet2018].
Below the intensity threshold the size of the surviving community maintained the pre-disturbance high taxonomic richness while the recruitment of pioneers, infrequent or absent before disturbance, increased both community taxonomic richness and evenness.
Beyond the intensity threshold, the disturbance decreased the taxonomic richness of surviving trees which was not offset by the enrichment of pioneers, so that the overall community taxonomic richness decreased according to the disturbance intensity [@Molino2001].
For community taxonomic evenness the disturbance impact was similar but slighter, as the evenness is less sensitive to the loss of rare species.
Taxonomic evenness rather represented the increasing dominance of pioneers that balanced the usual hyper-dominance of a few species in tropical forests below the intensity threshold, thus increasing community overall evenness up to the intensity threshold beyond which pioneers became in turn highly dominant and decreased the overall evenness [@Baraloto2012a].

>>>>

Dans ce chapitre nous nous intéresserons aux trajectoires de diversité et de composition à l'échelle de l'ensemble de la communauté.
La diversité des forêts tropicales est supposée entretenue naturellement par un régime régulier de perturbations entraînant un pic de diversité pour des perturbations d'intensité et de de fréquence moyennes.
En milieu tropical cependant cette théorie des perturbation intermédiaires reste débattue, ce qui questionne la résilience fonctionnelle et taxonomique des forêts tropicales.

@@

Les trajectoires de diversité ont mis en évidence des réponses taxonomique et fonctionnelle cycliques, permettant la restauration des caractéristiques d'avant perturbation.
Les trajectoires ont montré la résilience de la divergence taxonomique des communautés, due aux limites de dispersion des espèces maintenant les différences de composition initiales, et leur convergence fonctionnelle, due à des processus déterministes imposant une trajectoire fonctionnelle commune.
La théorie des perturbations intermédiaires a prédit correctement les trajectoires de diversité taxonomique, augmentant avec l'intensité de la perturbation jusqu'à un certain seuil.
En revanche la diversité fonctionnelle augmentaient après perturbation quelle qu'en soit l'intensité, l'impact de la perturbation étant atténué par l'importante redondance fonctionnelle des communautés.
Bien qu'effective, la restauration des caractéristiques taxonomiques et fonctionnelles des communautés restait inachevée après 30 ans, suggérant un temps de restauration long de plusieurs décennies.
La durée de la restauration s'est avérée déterminée par la restauration de la redondance fonctionnelle, montrant elle-même un ralentissement au cours du temps.

@@

Ariane MIRABEL ^1^ ^\*^, Bruno HERAULT ^2^, Eric MARCON ^1^
\newline

@@

Validations of the IDH though remain scarce in the long-term and mainly rely on the analysis of taxonomic richness [@Molino2001]. Taxonomic richness alone, however, gives limited or misleading information on forests recovery and functioning [@Chaudhary2016]. More ecological-meaningful analysis would couple richness with (i) evenness, that would reveal the changes in the species abundance distribution and thus the underlying ecological processes, and (ii) composition that is crucial for conservation issues [@Lavorel2002;@Bellwood2006]. Furthermore, a functional approach accounting for species biological attributes would directly link communities diversity, composition and redundancy to ecosystem functioning and to its environmental constraints [@Violle2007b;@Baraloto2012].
In that respect, the functional trait-based approach that focus on major traits related to species ecology and mediate species performance in a given environment was successfully adopted [@Diaz2005].
For instance, the functional approach revealed in tropical rainforests the deterministic processes entailing, after disturbance, a functional shift from a dominance of "conservative" slow-growing species dealing with scarce resources to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@Reich2014;@Herault2011].
This shift is translated into the trajectories of key functional traits related to resource acquisition (leaf and stem traits) and life-history traits (seed mass, maximum size) [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
Eventually a complete overview of community response to disturbance would encompass the changes in functional redundancy, that quantifies the amount of shared trait values among species [@Carmona2016].
The high functional redundancy of hyper-diverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines the resilience of communities after disturbance [@Elmqvist2003; @Diaz2005].

@@

The IDH well predicted well the disturbance impact on community taxonomic richness, enhanced until an intensity threshold (20-25% AGB loss), and to some extent on taxonomic evenness, somewhat decoupled from the disturbance intensity as already observed in the Guiana Shield [@Baraloto2012] and in Bornean tropical forests [@Cannon1998].
The disturbance intensity determined the balance in the community between pre-disturbance surviving trees and those recruited afterward.
The pool of true pioneer species specifically recruited after disturbance is restricted in the Guiana Shield to a few common genera (e.g. Cecropia spp., Vismia spp.) [@Guitet2018].
Below the intensity threshold the size of the surviving community maintained the pre-disturbance high taxonomic richness while the recruitment of pioneers, infrequent or absent before disturbance, increased both community taxonomic richness and evenness.
Beyond the intensity threshold, the disturbance decreased the taxonomic richness of surviving trees which was not offset by the enrichment of pioneers, so that the overall community taxonomic richness decreased according to the disturbance intensity [@Molino2001].
For community taxonomic evenness the disturbance impact was similar but slighter, as the evenness is less sensitive to the loss of rare species.
Taxonomic evenness rather represented the increasing dominance of pioneers that balanced the usual hyper-dominance of a few species in tropical forests below the intensity threshold, thus increasing community overall evenness up to the intensity threshold beyond which pioneers became in turn highly dominant and decreased the overall evenness [@Baraloto2012].

### 2018-09-12 05-RecruitmentDynamics.Rmd
discussion_commBH

La trajectoire des communautés après perturbation est déterminée par l'émergence de processus de recrutement déterministes favorisant les espèces héliophiles, remplaçant temporairement le recrutement stochastique propre aux communautés matures. 
Bien que résilientes les caractéristiques taxonomiques et fonctionnelles du recrutement se sont montrée impactées sur plusieurs décennies, incitant à la prudence quantd au maintien des caractéristiques initiales.

@@

Ariane MIRABEL ^1^ ^\*^, Eric Marcon ^1^, Bruno Hérault ^2^ \newline

@@

Determining the response of tropical forests to disturbance is key to predict their fate in the global changing context.
In the last decades, tropical forests experienced a wide range of disturbance, from radical land-use changes for agriculture or mining [@Dezecache2017a;@Dezecache2017b] to more insidious changes of communities structure, diversity and functioning following climatic changes [@Aubry-Kientz2015] or anthropogenic activities like selective logging [@Baraloto2012a]. 
In that respect a vast literature successfully modeled community response to disturbance in terms of tree growth, tree height and fluxes of carbon, water and nutrients [@Gourlet-Fleury2000;@Putz2012;@Piponiot2016;@Rutishauser2016].
Similar modeling approaches regarding forest diversity and composition remain hindered by the scarcity of long-term monitoring and by studies' restriction to common or commercial species imposed by forest huge biological diversity [@Sebbenn2008;@Vinson2015].
The template of community response to disturbance is set by recruitment processes that determine the new species joining the community.
Focusing on long-term recruitment trajectories therefore give valuable insight into post-disturbance recovery and hence into the adjustment of exploitation and conservation guidelines [@Diaz2005;@Schwartz2017].

@@

The taxonomic similarity between recruited trees and pre-disturbance forest was measured with the turnover metrics detailed in @Podani2013a.

@@

After disturbance the stochastic recruitment specific to undisturbed communities was progressively restored and drove community taxonomic and functional recovery. 
This confirmed previous results from the Paracou experiment, conducted 10 years [@Molino2001] and 20 years [@Baraloto2012a] after disturbance, where the early signs of the resilience of pre-disturbance taxonomic and functional composition recovery had been detected.

>>>>

La trajectoire des communautés après perturbation est déterminée par l'émergence de processus de recrutement déterministes favorisant les espèces héliophiles, remplaçant temporairement le recrutement stochastique propre aux communautés matures. 
Bien que résilientes les caractéristiques taxonomiques et fonctionnelles du recrutement se sont montrée impactées sur plusieurs décennies, incitant à la prudence quant au maintien des caractéristiques initiales.

@@

Ariane MIRABEL ^1^ ^\*^, Eric MARCON ^1^, Bruno HERAULT ^2^ \newline

@@

Determining the response of tropical forests to disturbance is key to predict their fate in the global changing context.
In the last decades, tropical forests experienced a wide range of disturbance, from radical land-use changes for agriculture or mining [@Dezecache2017a;@Dezecache2017b] to more insidious changes of communities structure, diversity and functioning following climatic changes [@Aubry-Kientz2015] or anthropogenic activities like selective logging [@Baraloto2012]. 
In that respect a vast literature successfully modeled community response to disturbance in terms of tree growth, tree height and fluxes of carbon, water and nutrients [@Gourlet-Fleury2000;@Putz2012;@Piponiot2016;@Rutishauser2016].
Similar modeling approaches regarding forest diversity and composition remain hindered by the scarcity of long-term monitoring and by studies' restriction to common or commercial species imposed by forest huge biological diversity [@Sebbenn2008;@Vinson2015].
The template of community response to disturbance is set by recruitment processes that determine the new species joining the community.
Focusing on long-term recruitment trajectories therefore give valuable insight into post-disturbance recovery and hence into the adjustment of exploitation and conservation guidelines [@Diaz2005;@Schwartz2017].

@@

The taxonomic similarity between recruited trees and pre-disturbance forest was measured with the turnover metrics detailed in @Podani2013.

@@

After disturbance the stochastic recruitment specific to undisturbed communities was progressively restored and drove community taxonomic and functional recovery. 
This confirmed previous results from the Paracou experiment, conducted 10 years [@Molino2001] and 20 years [@Baraloto2012] after disturbance, where the early signs of the resilience of pre-disturbance taxonomic and functional composition recovery had been detected.
